exp1_lik_by_trial

### 2021-06-04 comp_model_1019.Rmd
added report in papaja and softmax to BCH models

## First attempt : Q-Learning with state space consisting of last round play (deprecated) 
```{r}

# Q_learn_states <- function(par,data,return_value,gamma){
#   # Par[1] -> beta= inverse temperature parameter in softmax choice function
#   # Par[2] -> lambda = learning rate (one per game?)
#   # data : participant and ai choice data.
#   # return_value=c("-2loglik","likelihood_by_trial")
#   # Gamma is discount factor for future rewards 
#   # Returns Q-values per trial and predicts choice using softmax
#   beta <- par[1]
#   alpha <- par[2]
# 
#   #gamma <- 0.9 
# 
#   #Define matrix of state spaces for each game 
#   G1 <- expand.grid(c("R", "P", "S"),c("R", "P", "S"))
#   states_RPS <- paste0(G1$Var1,G1$Var2)
# 
#   G2 <- expand.grid(c("F", "W", "G"), c("F", "W", "G"))
#   states_FWG <- paste0(G2$Var1,G2$Var2)
# 
#   G3 <- expand.grid(c("1", "2", "3", "4", "5"), c("1", "2", "3", "4", "5"))
#   states_NUM <- paste0(G3$Var1,G3$Var2)
#   
#   Q_vals_RPS = matrix(-0.5,9,3)
#   dimnames(Q_vals_RPS) = list(states_RPS, c("R", "P", "S"))
# 
#   Q_vals_FWG = matrix(-0.5,9,3)
#   dimnames(Q_vals_FWG) = list(states_FWG, c("F", "W", "G"))
#   
#   Q_vals_NUM = matrix(-0.5,25,5)
#   dimnames(Q_vals_NUM) = list(states_NUM, c("1", "2", "3", "4", "5"))
#  
#   lik_hum <- matrix(0.0,nrow(data))
# 
#   for(t in 1:nrow(data)) {
#     t_game <- data[t,"game"]
#     if(t_game == "numbers") nopts <- 5 else nopts <- 3
#     
#     if(data[t,"round"] == 1) {
#       # first round is uniform prediction
#       Q_vals <- switch(as.character(t_game),rps=Q_vals_RPS,fwg = Q_vals_FWG, numbers = Q_vals_NUM)
#       state_vec <- switch(as.character(t_game),rps=states_RPS,fwg = states_FWG, numbers = states_NUM)
#       lik_hum[t] <- 1/nopts
#       # Randomly select prev_state and actions for first round
#       curr_state <- sample(state_vec, size = 1)
#       h_act <- sample(colnames(Q_vals), size =1)
#       ai_act <- sample(colnames(Q_vals), size =1)
#       reward <- 0 
#       
#     } else {
#       # Get past human action and associated reward 
#       h_act <- as.character(data[t-1,"h_action"])
#       ai_act <- as.character(data[t-1,"a_action"])
#       reward <- as.numeric(data[t-1,"score"])
#     }
#     
#     
#     # cat(curr_state, " This is the previous state", "\n")
#     new_state <- paste0(h_act,ai_act)
#     # cat(new_state,"This is new state","\n")
#     
#     # Q_learning: update rule (time  = t-1)
#     Q_vals[curr_state, h_act] <- Q_vals[curr_state, h_act] + alpha*( reward + gamma*max(Q_vals[new_state,]) - Q_vals[curr_state, h_act])
#     
#     # Assume human chooses action probabilistically using softmax on Q values
#     probs <- exp(Q_vals[new_state,]/beta)/sum(exp(Q_vals[new_state,]/beta))
#     #if (data[t,"round"] == 50) { cat(Q_vals,"\n") }
#     
#     # Get actual human action and compute likelihood
#     h_act <- as.character(data[t,"h_action"])
#     act_index <- match(h_act, colnames(Q_vals))
#     lik_hum[t] <- probs[[act_index]]
#       
#     # Update state
#     curr_state <- new_state 
#     #}
#   }
#   if(return_value == "-2loglik") {
#     ret <- -2*sum(log(lik_hum))
#     if(is.infinite(ret) || is.nan(ret)) {
#       return(1e+300)
#     } else {
#       return(ret)
#     }
#   }
#   if(return_value == "likelihood_by_trial") return(lik_hum)
#   
# }

@@

##  First attempt : Self-Tuning EWA with prior round play as state (Camerer & Ho 2007) 
# Initiates N(0) = 1 as in Camerer and Ho 1997 paper. 
# 
# EWA_states <- function(par,data,return_value){
# 
#   lambda <- par[1]
# 
#   #Define matrix of state spaces for each game 
#   G1 <- expand.grid(c("R", "P", "S"),c("R", "P", "S"))
#   states_RPS <- paste0(G1$Var1,G1$Var2)
# 
#   G2 <- expand.grid(c("F", "W", "G"), c("F", "W", "G"))
#   states_FWG <- paste0(G2$Var1,G2$Var2)
# 
#   G3 <- expand.grid(c("1", "2", "3", "4", "5"), c("1", "2", "3", "4", "5"))
#   states_NUM <- paste0(G3$Var1,G3$Var2)
#   
#   A_RPS = matrix(-0.5,9,3)
#   dimnames(A_RPS) = list(states_RPS, c("R", "P", "S"))
# 
#   A_FWG = matrix(-0.5,9,3)
#   dimnames(A_FWG) = list(states_FWG, c("F", "W", "G"))
#   
#   A_NUM = matrix(-0.5,25,5)
#   dimnames(A_NUM) = list(states_NUM, c("1", "2", "3", "4", "5"))
# 
#   
#   # # Define reward matrices from the perspective of the row player (human in our case)
#   reward_RPS <- t(matrix(c(0,-1,1,1,0,-1,-1,1,0),nrow=3,ncol=3))
#   dimnames(reward_RPS) = list(c("R", "P", "S"), c("R", "P", "S"))
#   
#   reward_FWG <- t(matrix(c(0,-1,1,1,0,-1,-1,1,0),nrow=3,ncol=3))
#   dimnames(reward_FWG) = list(c("F", "W", "G"), c("F", "W", "G"))
#   
#   reward_NUM <- t(matrix(c(0,-1,0,0,1,1,0,-1,0,0,0,1,0,-1,0,0,0,1,0,-1,-1,0,0,1,0),nrow=5, ncol=5))
#   dimnames(reward_NUM) = list(c("1", "2", "3", "4", "5"), c("1", "2", "3", "4", "5"))
#   
#   # Initiate likelihood by trial vector 
#   lik_hum <- matrix(0.0,nrow(data))
# 
#   for(t in 1:nrow(data)) {
#     t_game <- data[t,"game"]
#     if(t_game == "numbers") nopts <- 5 else nopts <- 3
#     if(data[t,"round"] == 1) {
#       Att <- switch(as.character(t_game),rps=A_RPS,fwg = A_FWG, numbers = A_NUM)
#       state_vec <- switch(as.character(t_game), rps=states_RPS, fwg = states_FWG, numbers = states_NUM)
#       reward <- switch(as.character(t_game), rps=reward_RPS, fwg = reward_FWG, numbers = reward_NUM)
#       game_data <- subset(data, game == t_game)
#       
#       # first round is uniform prediction
#       lik_hum[t] <- 1/nopts
#       N <- 1.0
# 
#       R_t <- rep(0.0,nopts)
#       H_t <- R_t
#       
#       # Randomly select prev_state and actions for first round
#       curr_state <- sample(state_vec, size = 1)
#       h_act_prev <- sample(colnames(Att), size =1)
#       a_act_prev <- sample(colnames(Att), size =1)
#       #reward <- 0 
#       
#     } else {
#       
#       indx <- data[t,"round"]
#       state_indx <- match(curr_state,state_vec)
#       #cat("this is the current state index", state_indx, "\n")
#       
#       # Get reward and past human action
#       h_act_prev <- as.character(game_data[indx-1,"h_action"])
#       a_act_prev <- as.character(game_data[indx-1,"a_action"])
#       new_state <- paste0(h_act_prev,a_act_prev)
# 
#       # Estimate R(t) (recent history) then H(t) (history) then phi(t) (change detector)
#       R_t <- as.numeric(colnames(Att) == a_act_prev)
#       #cat("This is R(t)",R_t,"\n")
#       
#       H_t <- (H_t*(indx-2) + R_t)/(indx-1)
#       #cat("this is H(t)",H_t,"\n")
#       
#       Phi_t <- 1 - 0.5*sum((H_t-R_t)^2)
#       #cat("This is Phit(t)",Phi_t,"\t")
#       
#       # Estimate vector Delta(t)
#       delta_t <- as.numeric(reward[,a_act_prev] >= as.numeric(game_data[indx-1,"score"]))
#       #cat("this is delta(t)",delta_t,"\n")
#       
#       for (i in 1:nopts) {
#         action <- as.character(colnames(Att)[i])
# 
#         #cat("this is current indexed action",action,"\n")
#         #cat("this is previous human action",h_act_prev,"\n")
#         
#         # Attraction vector update rule
#         Att[state_indx,i] <- (Phi_t*N*Att[state_indx,i]  + ( delta_t[[i]] + (1- delta_t[[i]])*(action == h_act_prev))*as.numeric(reward[action,a_act_prev])) / (Phi_t*N + 1)
#       }
#       
#       #Update the value of N 
#       N <- Phi_t*N + 1
# 
#       # Assume human chooses action probabilistically using softmax on Attraction values
#       probs <- exp(lambda*Att[new_state,])/sum(exp(lambda*Att[new_state,]))
#       
#       # Get actual human action and compute likelihood
#       h_act <- as.character(game_data[indx,"h_action"])
#       act_index <- match(h_act, colnames(Att))
#       lik_hum[t] <- probs[[act_index]]
#       
#       curr_state <- new_state
#       
#       # if (indx == 50) {
#       #   cat("this is the attraction matrix at last round of", "\n")
#       #   print(Att)
#       # }
#     }
#   }
#   
#  if(return_value == "-2loglik") {
#     ret <- -2*sum(log(lik_hum))
#     if(is.infinite(ret) || is.nan(ret)) {
#       return(1e+300)
#     } else {
#       return(ret)
#     }
#   }
#   if(return_value == "likelihood_by_trial") return(lik_hum)
# }

@@

## Putting results together 
```{r}
# load("QL_modelling.RData")
# load("QL_states_modelling.RData")
# load("EWA_modelling.RData")
# load("EWA_self_modelling.RData")
# load("MBM_modelling.RData")
# load("ST_EWA_STATES_modelling.RData")

@@

                         # Bayesian updating with/without transfer
                         "Bayes_Tr_BIC" = exp1_Bayes_game_Tr[[id]]$optim$bestval+ 1*log(150),
                         "Bayes_No_Tr_BIC" = exp1_Bayes_no_Tr[[id]]$optim$bestval + 1*log(150),
                         # Theta is the parameter governing AI stochasticity. Truth is 0.9
                         "theta_transfer" = exp1_Bayes_game_Tr[[id]]$optim$bestmem[[1]],
                         "theta_no_transfer" = exp1_Bayes_no_Tr[[id]]$optim$bestmem[[1]],

>>>>

@@

@@

## Putting results together 
```{r}
load("QL_modelling.RData")
load("QL_states_modelling.RData")
load("EWA_modelling.RData")
load("EWA_self_modelling.RData")
load("MBM_modelling.RData")
load("ST_EWA_STATES_modelling.RData")

@@

                         # Bayesian updating with/without transfer 
                         # "Bayes_Tr_BIC" = exp1_Bayes_game_Tr[[id]]$optim$bestval+ 1*log(150),
                         # "Bayes_No_Tr_BIC" = exp1_Bayes_no_Tr[[id]]$optim$bestval + 1*log(150),
                         # # Theta is the parameter governing AI stochasticity. Truth is 0.9
                         # "theta_transfer" = exp1_Bayes_game_Tr[[id]]$optim$bestmem[[1]],
                         # "theta_no_transfer" = exp1_Bayes_no_Tr[[id]]$optim$bestmem[[1]],
                         #

                         # BCH with/without transfer (With Softmax) 
                         "Bayes_Tr_BIC" = use_softmax_tr[[id]]$optim$bestval+ 2*log(150),
                         "Bayes_No_Tr_BIC" = use_softmax_NT[[id]]$optim$bestval+ 2*log(150),

                         # Theta is the parameter governing AI stochasticity. Truth is 0.9
                         "theta_transfer" = use_softmax_tr[[id]]$optim$bestmem[[1]],
                         "lambda_transfer" = use_softmax_tr[[id]]$optim$bestmem[[2]],

                         "theta_no_transfer" = use_softmax_NT[[id]]$optim$bestmem[[1]],
                         "lambda_no_transfer" = use_softmax_NT[[id]]$optim$bestmem[[2]],

### 2021-06-04 exp1_lik_by_trial.Rmd
added report in papaja and softmax to BCH models

  #Bayes model no transfer likelihoods
  theta_NT <- unlist(filter(optim_results, ID == id)%>% dplyr::select(theta_no_transfer))
  tdat$Bayes_NT_lik <- exp1_Bayes_model_LL(theta_NT ,data = tdat, generalize = "no", "likelihood_by_trial")

  # Bayes model with Transfer likelihood
  theta_Tr <- unlist(filter(optim_results, ID == id)%>% dplyr::select(theta_transfer))
  tdat$Bayes_Tr_lik <- exp1_Bayes_model_LL(theta_Tr ,data = tdat, generalize = "game", "likelihood_by_trial")

@@

mean_lik <- exp1_lik_by_trial %>%
  select(human_id,condition,game, round,QLS_lik,Bayes_NT_lik,Bayes_Tr_lik) %>%
  group_by(game, condition, round)  %>%
  dplyr::summarise(mean_QLS_lik = mean(QLS_lik, na.rm = TRUE),
            mean_NT_lik = mean(Bayes_NT_lik, na.rm = TRUE),
            mean_Tr_lik = mean(Bayes_Tr_lik, na.rm = TRUE))

>>>>

  #Bayes model no transfer likelihoods
  pars_BCH_NT <- unlist(filter(optim_results, ID == id)%>% dplyr::select(theta_no_transfer,lambda_no_transfer))
  tdat$Bayes_NT_lik <- exp1_Bayes_model_LL(pars_BCH_NT ,data = tdat, generalize = "no", softmax = TRUE, "likelihood_by_trial")

  # Bayes model with Transfer likelihood
  pars_BCH_Tr <- unlist(filter(optim_results, ID == id)%>% dplyr::select(theta_transfer,lambda_transfer))
  tdat$Bayes_Tr_lik <- exp1_Bayes_model_LL(pars_BCH_Tr,data = tdat, generalize = "game", softmax = TRUE, "likelihood_by_trial")

@@

mean_lik <- exp1_lik_by_trial %>%
  dplyr::select(human_id,condition,game, round,QLS_lik,Bayes_NT_lik,Bayes_Tr_lik) %>%
  group_by(game, condition, round)  %>%
  dplyr::summarise(mean_QLS_lik = mean(QLS_lik, na.rm = TRUE),
            mean_NT_lik = mean(Bayes_NT_lik, na.rm = TRUE),
            mean_Tr_lik = mean(Bayes_Tr_lik, na.rm = TRUE))
